Bootstrapping Multi-atlas Hippocampal Segmentation with \mb

### 2013-08-03 paper.Rnw
Existing work table updated, last few fixes.
In particular, I added a blurb about how MP did QC. Also, added in a short
comparison to other existing work (referencing the table) and referenced
the table when justifying our low overlap result

Automated segmentation techniques that are reliable, objective, and reproducible
are a necessary alternative to manual segmentation.  In the case of classical
model-based segmentation methods \citep{Haller1997,Csernansky1998}, an MRI atlas
that was previously manually labelled by an expert rater is matched to target
images using nonlinear registration methods.  The resulting nonlinear
transformation is applied to the manual labels (ie. {\em label propagation}) to
move them into the target image space.  While this methodology has been used
successfully in several contexts
\citep{Chakravarty2008,Chakravarty2009,Collins1995,Haller1997}, it is limited
in accuracy by the introduction of errors due inaccuracies in the nonlinear
transformation itself, partial volume effects in label resampling, and
irreconcilable differences between the neuroanatomy represented within the atlas
and target images.

@@

More recently, many groups have been using multiple atlases to improve overall
segmentation accuracy (ie. multi-atlas segmentation) over model-based approaches
\citep{Heckemann2006,Heckemann2011,Collins2010,Lotjonen2010,Aljabar2009,Leung2010,Wolz2010}.
Each atlas image is registered to a target image, and label propagation is
performed to produce several labellings of the target image (one from each
atlas). A {\it label fusion} technique, such as voxel-wise voting, is used to
merge these labels into a definitive segmentation for the target.  In addition,
weighted voting procedures that use {\em atlas selection} techniques are often
used to exclude atlases from label fusion that are dissimilar to a target image
in order to reduce error from unrepresentative anatomy \citep{Aljabar2009}.
This involves the selection of a subset of atlases using a similarity metric
such a cross-correlation \citep{Aljabar2009} or normalized mutual information.
Such selection has the added benefit of significantly reducing the number of
nonlinear registrations.  For example Collins and Pruessner \citep{Collins2010}
demonstrated that only 14 atlases, selected based on highest similarity between
medial temporal lobe neuroanatomy as evaluated by normalized mutual information
\citep{Studholme1999} from a library of 80 atlases, were required to achieve
accurate segmentations of the hippocampus.  Several methods have been explored
for label fusion including the STAPLE algorithm \citep{Warfield2004} that
computes a probabilistic segmentation using an expectation maximization
framework from an set of competing segmentations; or others where a subset of
segmentations can be estimated using metrics such as the sum of squared
differences in the regions of interest to be segmented \citep{Coupe2011}.

@@

There are few methods that have attempted to perform multi-atlas segmentation
with a limited number of input atlases. The LEAP algorithm is an elegant
modification to the basic multi-atlas strategy \citep{Wolz2010} in which the
atlas library is grown, beginning with a set of manually labelled atlases, and
successively incorporating unlabelled target images after themselves being
labelling using multi-atlas techniques. The sequence in which target images are
labelled is chosen so that the similarity between the atlas images and the
target images is minimised at each step, effectively allowing for deformations
between very dissimilar images to be broken up into sequences of smaller
deformations.  Although \citet{Wolz2010} begin with an atlas library of 30 MR
images, this method could theoretically work using a much smaller atlas library.
In their validation, LEAP was used to segment the whole hippocampus in the ADNI1
baseline dataset, achieving a mean Dice score of 0.85 with manual segmentations.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% ADNI1 X-Validation Dataset Demographics Table
%
<<ADNI-Xval-demographics, echo=F, results="asis",cache=FALSE>>=
adni_69_RIDs = c(295,413,619,685,729,782,938,954,1018,1155,907,981,222,324,448,
                 546,553,572,602,610,814,1341,675,681,731,1130,41,68,70,101,
                 249,293,316,414,698,1206,1222,1339,751,1030,3,5,10,16,22,53,
                 168,183,241,1205,328,1095,991,213,159,337,343,642,753,1109,
                 1217,29,56,1188,1293,149,382,431,1202)
tab <- summary(DX.bl ~ AGE + PTGENDER + PTEDUCAT + CDRSB + ADAS13 +
               MMSE, data = adnimerge, subset = VISCODE == 'bl' & RID %in%
               adni_69_RIDs, method = "reverse", test=FALSE, overall=TRUE)
caption = paste(
  "\\textbf{ADNI-1 cross-validation subset demographics.}", 
  "Clinical Dementia Rating-Sum of Boxes (CDR-SB), Alzheimer's Disease",
  "Assessment Scale (ADAS), Mini–Mental State Examination (MMSE).", sep=" ") 
latex(tab, file="", size="scriptsize",landscape=FALSE, ctable=FALSE,
      caption=caption, caption.loc = c("top"),
      title="tab:ADNI1-xvalidation-demographics")
@
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

The Dice similarity coefficient (DSC) assesses the agreement between two
segmentations. It is one of the most widely used measures of segmentation
performance, and we use it as the basis of comparison in this experiment.
Additionally, we report the Jaccard index, another commonly used similarity
measure:

@@

 \[\text{Jaccard (J)} = \frac{|A \cap B|}{|A \cup B|} = \frac{DSC}{(2-DSC)}\]

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Subfield Validation Template library demographics (healthy controls)
%                                     
<<WAval-healthy-demographics, echo=F, results="asis",cache=FALSE>>=
master = read.csv('data/DT_study.csv')
tab <- summary(~Age + Sex + Education + Handedness, 
               data = master, method="reverse", na.rm = TRUE, test=FALSE)
caption = paste(
  "\\textbf{ADNI-1 cross-validation subset demographics.}", 
  "Clinical Dementia Rating-Sum of Boxes (CDR-SB), Alzheimer's Disease",
  "Assessment Scale (ADAS), Mini–Mental State Examination (MMSE).", sep=" ") 
latex(tab, file="", size="scriptsize",landscape=FALSE, ctable=FALSE,
      caption=caption, caption.loc = c("top"),
      title="tab:ADNI1-xvalidation-demographics")
@
\begin{table*}[!tbp]
\scriptsize
\caption{\textbf{Subfield validation template library demographics.}
\label{tab:WAval-healthy-demographics}}
\begin{center}
\begin{tabular}{l | c c c}
\hline 
Parameters        & Stage 1 & Stage 2 & Stage 3  \tabularnewline
\hline 
Model Blur (FWHM) &   8     &    8    &   4      \tabularnewline
Input Blur (FWHM) &   8     &    8    &   4      \tabularnewline
Iterations        &   30    &    30   &   10     \tabularnewline
Step              &   8x8x8 &  4x4x4  &   2x2x2  \tabularnewline
Sub-Lattice       &   6     &    6    &   6      \tabularnewline
Lattice Diameter  &24x24x24 &12x12x12 & 6x6x6    \tabularnewline
\hline
\end{tabular}
\end{center}
\end{table*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ADNI1-XVAL: show variability of MAGeT over parameters
\begin{figure}
<<ADN1-Xval-variability, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
ants = subset(all_data_mean, 
                     reg_method=="ANTS" & 
                     method.mb=="Majority Vote" &
                    templates.mb %in% seq(1,20,2))

ants.stats = ddply(ants, c("subject", "atlases","templates.mb"), 
                   function (df) {
                     data.frame(
                       MA.sd  =  sd(df$k.ma),
                       MB.sd  =  sd(df$k.mb),
                       MA.var = var(df$k.ma),
                       MB.var = var(df$k.mb)
                       )
                   })

stats.tests = ddply(data.stats, c("atlases","templates.mb"), function (df) {
  t.var = t.test(df$MA.var, df$MB.var)
  t.sd  = t.test(df$MA.sd , df$MB.sd)
  diff_mean_var = mean(df$MA.var - df$MB.var)
  diff_mean_sd  = mean(df$MA.sd -  df$MB.sd)

@@

ggplot(subset(stats.tests, direction > 0 & stat == "Variance"), 
  aes(x=templates.mb, y=p.value, colour=as.factor(atlases))) + 
  #facet_grid( . ~ stat) + 
  geom_line(size=1) +
  geom_point(size=3) + 
  geom_hline(aes(yintercept=0.05, alpha=0.5), linetype='dashed') + 
  geom_hline(aes(yintercept=0.01, alpha=0.5), linetype='dashed') + 
  scale_x_continuous(breaks=seq(1,20,by=2)) + 
  scale_y_log10(limits=c(0.001,1), breaks=c(0.01,0.05,0.1,1.0)) + 
  xlab( "Number of Templates" ) + 
  ylab( "Difference of variance during validation (p-value)" ) +
  scale_colour_hue(name="Atlases") +
  scale_alpha_continuous(guide='none') + 
  scale_linetype_discrete(guide='none') + 
  scale_size_continuous(guide='none') + 
  theme(legend.direction = "horizontal", legend.position = "bottom")
@
  \caption{{\bf Difference in variability of \mb vs. multi-atlas segmentation 
  accuracy.}
  Variance of segmentation accuracy between \mb and multi-atlas segmentation 
  is computed for each subject across all ten rounds of validation. Shown on the 
  y-axis (scaled logarithmically) is the  p-value resulting from a t-test 
  comparing the distribution of variances at each parameter setting 
  (atlas/template library size). Only points where \mb mean variability is lower 
  than multi-atlas are shown.}
  \label{ADNI1-xval-variability}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Based on the results from the ADNI1 Cross-Validation experiment, in this
experiment \mb was configured with a template library of 21 randomly chosen
subject images (7 from each disease class) and used majority vote label fusion.
The entire \adnidataset dataset was segmented by \mb and we now compare the
resulting volumes with those obtained by manual segmentation (SNT), and other
automated segmentation techniques (MAPER, FreeSurfer, and FSL).  Table
\ref{tab:ADNI1-1.5T-Complete-1Yr-Dataset-Method-Totals} shows the total count of
segmentations available, including a count of those which have failed a quality
control inspection. Quality control for FreeSurfer segmentations was provided along
with the segmentations themselves from ADNI \cite{Hartig2010}. One author (MP) 
inspected

Only those images which had segmentations from every method
are included in the following analysis (a total of
\Sexpr{length(means.complete$RID)} images).

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ADNI1 1yr Complete Dataset Package Totals & Failures
%                        
<<ADNI-seg-package-totals, echo=F, results="asis",cache=FALSE>>=
package_totals$Total <- NULL
latex(package_totals, file="", size="scriptsize",landscape=FALSE, ctable=TRUE,
    caption="\\adnidataset dataset segmentation totals and QC failures.", 
    caption.loc = c("top"), rowname=NULL, 
    label="tab:ADNI1-1.5T-Complete-1Yr-Dataset-Method-Totals")
@
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                Discussion
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Tabulation of segmentation accuracies of existing methods (ADNI)
%
<<ADNI1-literature, echo=F, results="asis",cache=FALSE>>=
caption = paste(
  "Survey of automated segmentation accuracy of the ADNI dataset",
  sep=" ")

tab = read.csv("data/ADNI-existing-work.csv")
tab$Notes <- NULL

adni_lit = subset(tab, Dataset=="ADNI")
adni_lit$Dataset <- NULL
other_lit = subset(tab, Dataset!="ADNI")
other_lit$Dataset <- NULL

latex(adni_lit, file="", 
      size="scriptsize",
      caption=caption, 
      rowname=NULL,
      col.just=c("p{1.5in}l","p{0.5in}","p{1.3in}","p{1.5in}"),
      title="")
@
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Throughout the cross validation in Experiment 1 (10-fold cross validation in the
\adnidataset dataset subsample) we find that two parameter choices improve 
segmentation accuracy: increasing the number of atlases, and the number of
templates. However, after setting the parameters to 5 atlases and 15 templates
there are diminishing returns with respect to this improvement. We were
somewhat surprised to find that cross-correlation and normalized mutual
information based weighted voting did little to improve segmentation accuracy.
This suggests that using an intermediate template library generated by accurate
nonlinear registration methods such as \ants \citep{Avants2008} sufficiently filters
sources of error such as nonlinear registration error and partial volume effects
that arise through the use of nearest neighbour resampling. From our previous
work on the \mb algorithm we have shown that these increases in error
are not simply a smoothing or averaging effect \citep{MallarChakravarty2012}.

@@

Another concern is the moderate-to-small improvement observed in \mb in
comparison to multi-atlas segmentation when using the same number of atlases.
The actual benefit in using \mb is consistency of the labelling regardless of
atlas or template choice. This is an important consideration that few have
touched on previously. The 10-fold Monte Carlo leave-one-out cross-validation that we present
in Experiment 1 is amongst one of the most stringent performed in the
multi-atlas/segmentation literature.  To the best of our knowledge other groups
have only ever performed cross-validation \citep{?}.  This suggests that
our results are reflective of a true average over the choice of parameter
settings and are independent of atlas or template choice (provided the input
atlases are properly segmented).

% Validation in other works
% Chupin 2009: LOOCV, 1-fold validation
% Heckemann 2011: QC of segmentations, comparison across field strengths, 
%                 agreement between two different atlas sets
% Coupe 2012: LOOCV
% Lotjonen 2010: LOOCV, 1-fold validation
% Leung 2010: 1-fold, QC, classification
% Wolz 2010: 1-fold over different leap settings, classification
% Collins 2010: 1-fold validation, QC
%

>>>>

Automated segmentation techniques that are reliable, objective, and reproducible
are a necessary alternative to manual segmentation.  In the case of classical
model-based segmentation methods \citep{Haller1997,Csernansky1998}, an MRI atlas
that was previously manually labelled by an expert rater is matched to target
images using nonlinear registration methods.  The resulting nonlinear
transformation is applied to the manual labels (i.e. {\em label propagation}) to
move them into the target image space.  While this methodology has been used
successfully in several contexts
\citep{Chakravarty2008,Chakravarty2009,Collins1995,Haller1997}, it is limited
in accuracy by the introduction of errors due inaccuracies in the nonlinear
transformation itself, partial volume effects in label resampling, and
irreconcilable differences between the neuroanatomy represented within the atlas
and target images.

@@

More recently, many groups have been using multiple atlases to improve overall
segmentation accuracy (i.e. multi-atlas segmentation) over model-based approaches
\citep{Heckemann2006,Heckemann2011,Collins2010,Lotjonen2010,Aljabar2009,Leung2010,Wolz2010}.
Each atlas image is registered to a target image, and label propagation is
performed to produce several labellings of the target image (one from each
atlas). A {\it label fusion} technique, such as voxel-wise voting, is used to
merge these labels into a definitive segmentation for the target.  In addition,
weighted voting procedures that use {\em atlas selection} techniques are often
used to exclude atlases from label fusion that are dissimilar to a target image
in order to reduce error from unrepresentative anatomy \citep{Aljabar2009}.
This involves the selection of a subset of atlases using a similarity metric
such a cross-correlation \citep{Aljabar2009} or normalized mutual information.
Such selection has the added benefit of significantly reducing the number of
nonlinear registrations.  For example \citet{Collins2010}
demonstrated that only 14 atlases, selected based on highest similarity between
medial temporal lobe neuroanatomy as evaluated by normalized mutual information
\citep{Studholme1999} from a library of 80 atlases, were required to achieve
accurate segmentations of the hippocampus.  Several methods have been explored
for label fusion including the STAPLE algorithm \citep{Warfield2004} that
computes a probabilistic segmentation using an expectation maximization
framework from an set of competing segmentations; or others where a subset of
segmentations can be estimated using metrics such as the sum of squared
differences in the regions of interest to be segmented \citep{Coupe2011}.

@@

There are few methods that have attempted to perform multi-atlas segmentation
with a limited number of input atlases. The LEAP algorithm is an elegant
modification to the basic multi-atlas strategy \citep{Wolz2010} in which the
atlas library is grown, beginning with a set of manually labelled atlases, and
successively incorporating unlabelled target images after themselves being
labelling using multi-atlas techniques. The sequence in which target images are
labelled is chosen so that the similarity between the atlas images and the
target images is minimised at each step, effectively allowing for deformations
between very dissimilar images to be broken up into sequences of smaller
deformations.  Although \citet{Wolz2010} begin with an atlas library of 30 MR
images, this method could theoretically work using a much smaller atlas library.
In their validation, LEAP was used to segment the whole hippocampus in the
ADNI-1 baseline dataset, achieving a mean Dice score of 0.85 with manual
segmentations.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% ADNI1 X-Validation Dataset Demographics Table
%
<<ADNI-Xval-demographics, echo=F, results="asis",cache=FALSE>>=
adni_69_RIDs = c(295,413,619,685,729,782,938,954,1018,1155,907,981,222,324,448,
                 546,553,572,602,610,814,1341,675,681,731,1130,41,68,70,101,
                 249,293,316,414,698,1206,1222,1339,751,1030,3,5,10,16,22,53,
                 168,183,241,1205,328,1095,991,213,159,337,343,642,753,1109,
                 1217,29,56,1188,1293,149,382,431,1202)
tab <- summary(DX.bl ~ AGE + PTGENDER + PTEDUCAT + CDRSB + ADAS13 +
               MMSE, data = adnimerge, subset = VISCODE == 'bl' & RID %in%
               adni_69_RIDs, method = "reverse", test=FALSE, overall=TRUE)
caption = paste(
  "\\textbf{ADNI-1 cross-validation subset demographics.}", 
  "Clinical Dementia Rating-Sum of Boxes (CDR-SB), Alzheimer's Disease",
  "Assessment Scale (ADAS), Mini–Mental State Examination (MMSE).", sep=" ") 
latex(tab, file="", size="scriptsize",landscape=FALSE, ctable=FALSE,
      caption=caption, caption.loc = c("top"),
      label="tab:ADNI1-xvalidation-demographics")
@
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

The Dice similarity coefficient (DSC) assesses the agreement between two
segmentations. It is one of the most widely used measures of segmentation
performance, and we use it as the basis of comparison in this experiment.
%Additionally, we report the Jaccard index, another commonly used similarity
%measure:

@@

% \[\text{Jaccard (J)} = \frac{|A \cap B|}{|A \cup B|} = \frac{DSC}{(2-DSC)}\]

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Subfield Validation Template library demographics (healthy controls)
%                                     
<<WAval-healthy-demographics, echo=F, results="asis",cache=FALSE>>=
master = read.csv('data/DT_study.csv')
tab <- summary(~Age + Sex + Education + Handedness, 
               data = master, method="reverse", na.rm = TRUE, test=FALSE)
caption = paste(
  "\\textbf{ADNI-1 cross-validation subset demographics.}", 
  "Clinical Dementia Rating-Sum of Boxes (CDR-SB), Alzheimer's Disease",
  "Assessment Scale (ADAS), Mini–Mental State Examination (MMSE).", sep=" ") 
latex(tab, file="", size="scriptsize",landscape=FALSE, ctable=FALSE,
      caption=caption, caption.loc = c("top"),
      title="tab:WAval-healthy-demographics")
@
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ADNI1-XVAL: show variability of MAGeT over parameters
\begin{figure}
<<ADN1-Xval-variability, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
ants = subset(all_data, 
                     reg_method=="ANTS" & 
                     method.mb=="Majority Vote" &
                    templates.mb %in% seq(1,20,2))

ants.stats = ddply(ants, c("subject", "label", "atlases","templates.mb"), 
                   function (df) {
                     data.frame(
                       MA.sd  =  sd(df$k.ma),
                       MB.sd  =  sd(df$k.mb),
                       MA.var = var(df$k.ma),
                       MB.var = var(df$k.mb)
                       )
                   })

stats.tests = ddply(ants.stats, c("atlases","templates.mb"), function (df) {
  t.var = t.test(df$MA.var, df$MB.var)
  t.sd  = t.test(df$MA.sd , df$MB.sd)
  diff_mean_var = mean(df$MA.var - df$MB.var)
  diff_mean_sd  = mean(df$MA.sd -  df$MB.sd)

@@

ggplot(subset(stats.tests, direction > 0 & stat == "Variance"), 
  aes(x=templates.mb, y=p.value, colour=as.factor(atlases))) + 
  #facet_grid( . ~ stat) + 
  geom_line(size=1) +
  geom_point(size=3) + 
  geom_hline(aes(yintercept=0.05, alpha=0.5), linetype='dashed') + 
  geom_hline(aes(yintercept=0.01, alpha=0.5), linetype='dashed') + 
  scale_x_continuous(breaks=seq(1,20,by=2)) + 
  scale_y_log10(limits=c(0.001,1), breaks=c(0.01,0.05,0.1,1.0)) + 
  xlab( "Number of Templates" ) + 
  ylab( "Difference of variance during validation (p-value)") +
  scale_colour_hue(name="Atlases") +
  scale_alpha_continuous(guide='none') + 
  scale_linetype_discrete(guide='none') + 
  scale_size_continuous(guide='none') + 
  theme(legend.direction = "horizontal", legend.position = "bottom")
@
  \caption{{\bf Difference in variability of \mb vs. multi-atlas segmentation 
  accuracy.}
  Variance of segmentation accuracy between \mb and multi-atlas segmentation 
  is computed for each subject across all ten rounds of validation. Shown on the 
  y-axis (scaled logarithmically) is the  p-value resulting from a t-test 
  comparing the distribution of variances at each parameter setting 
  (atlas/template library size). Only points where \mb mean variability is lower 
  than multi-atlas are shown.}
  \label{ADNI1-xval-variability}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Based on the results from the ADNI1 Cross-Validation experiment, in this
experiment \mb was configured with a template library of 21 randomly chosen
subject images (7 from each disease class) and used majority vote label fusion.
The entire \adnidataset dataset was segmented by \mb and we now compare the
resulting volumes with those obtained by manual segmentation (SNT), and other
automated segmentation techniques (MAPER, FreeSurfer, and FSL).  Table
\ref{tab:ADNI1-1.5T-Complete-1Yr-Dataset-Method-Totals} shows the total count of
segmentations available, including a count of those which have failed a quality
control inspection. Quality control for FreeSurfer segmentations was provided along
with the segmentations from ADNI \cite{Hartig2010}. One of the authors (MP) 
performed visual quality inspection for MAGeT and FSL segmentations using similar 
quality control guidelines (if either hippocampus was under or over segmented
by greater than 10mm in three or more slices then the segmentation did not pass).

Only those images which had quality segmentations from each method are 
included in the following analysis (a total of \Sexpr{length(means.complete$RID)}
images).

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ADNI1 1yr Complete Dataset Package Totals & Failures
%                        
<<ADNI-seg-package-totals, echo=F, results="asis",cache=FALSE>>=
package_totals$Total <- NULL
latex(package_totals, file="", size="scriptsize",landscape=FALSE, ctable=TRUE,
    caption="\\textbf{Number of segmented images and quality control failures of 
             \\adnidataset dataset by method.}", 
    caption.loc = c("top"), rowname=NULL, 
    label="tab:ADNI1-1.5T-Complete-1Yr-Dataset-Method-Totals")
@
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                Discussion
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Tabulation of segmentation accuracies of existing methods (ADNI)
%
<<ADNI1-literature, echo=F, results="asis",cache=FALSE>>=
# caption = 
#   "\\textbf{Automated segmentation accuracy (overlap with manual labels) of the 
#    ADNI dataset.} For each method, the number of manually labelled atlases used for 
#    training, the best Dice's overlap measure, the disease classes measured, and 
#    the validation procedure are shown. LOOCV = Leave-one-out cross-validation. 
#    Some studies of automated segmentation of ADNI images are excluded because they 
#    do not not provide overlap measures for the hippocampus \\cite{Heckemann2011, Chupin2009}.
#   "
# 
# tab = read.csv("data/ADNI-existing-work.csv")
# tab$Notes <- NULL
# 
# adni_lit = subset(tab, Dataset=="ADNI")
# adni_lit$Dataset <- NULL
# 
# 
# latex(adni_lit, #file="", 
#       size="scriptsize",
#       caption=caption, 
#       rowname=NULL,
#       col.just=c("p{2in}","p{1in}","p{1in}","p{1in}","p{1in}"),
#       title="")
@
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Tabulation of segmentation accuracies of existing methods (ADNI)
%
%
<<mb-best-kappa, echo=F, cache=FALSE>>=
ants.mean = subset(all_data_mean, reg_method=="ANTS" & method.mb=="Majority Vote" &      templates.mb == 19 & atlases == 9)
mb.best.kappa = mean(ants.mean$k.mb)
@

\begin{table}[!tbp]
\scriptsize
\caption{\textbf{Automated segmentation accuracy (overlap with manual labels) of the 
   ADNI dataset.} For each method, the number of manually labelled atlases used for 
   training, the best Dice's overlap measure, the disease classes measured, and 
   the validation procedure are shown. Unless specified, validation datasets are
   composed equally of subjects diagnosed with Alzheimer's Disease (AD), Mild 
   Cognitive Impairment (MCI), and Cognitively Normal (CN). LOOCV = Leave-one-out 
   cross-validation. 
   Some studies of automated segmentation of ADNI images are excluded because they 
   do not not provide overlap measures for the hippocampus \cite{Heckemann2011,Chupin2009}.
  \label{tab:other-methods}} 
\begin{center}
\begin{tabular}{ l  c  p{1in}  l  p{2in}  }
\hline\hline
Method & Atlases & DSC & Reference & Validation \tabularnewline
\hline
\mb                     &9 & \Sexpr{round(mb.best.kappa, digits=3)} & & 10 rounds of Monte Carlo CV on a pool of 69 subjects \tabularnewline
LEAP                    &30&0.848                 &\cite{Wolz2010}  &Segmentation of 60 subjects\tabularnewline
ACM (AdaBoost-based)    &21&0.862                 &\cite{Morra2008} &LOOCV on atlases\tabularnewline
Patch-based label fusion&16&0.883 (CN) \newline 0.838 (AD)&\cite{Coupe2011e}&LOOCV on atlases\tabularnewline
Multi-atlas             &30&0.885                 &\cite{Lotjonen2010}&Segmentation of 60 subjects\tabularnewline
Multi-atlas + weighted fusion&20&0.898 (CN) \newline 0.798 (left HC, MCI) &\cite{Wang2011}  &10 rounds of Monte Carlo CV on 20 subjects, pool of 139 (CN/MCI)\tabularnewline
Multi-atlas (MAPS)    &55&0.890                   &\cite{Leung2010}&Segmentation of 30 subjects (10 AD, MCI, and CN)\tabularnewline
\hline
\end{tabular}
\end{center}
\end{table}

@@

Throughout the cross validation in Experiment 1 (10-fold Monte Carlo cross 
validation in the
\adnidataset dataset subsample) we find that two parameter choices improve 
segmentation accuracy: increasing the number of atlases, and the number of
templates. However, after setting the parameters to 5 atlases and 15 templates
there are diminishing returns with respect to this improvement. We were
somewhat surprised to find that cross-correlation and normalized mutual
information based weighted voting did little to improve segmentation accuracy.
This suggests that using an intermediate template library generated by accurate
nonlinear registration methods such as \ants \citep{Avants2008} sufficiently filters
sources of error such as nonlinear registration error and partial volume effects
that arise through the use of nearest neighbour resampling. From our previous
work on the \mb algorithm we have shown that these increases in error
are not simply a smoothing or averaging effect \citep{MallarChakravarty2012}.

@@

Another concern is the moderate-to-small improvement observed in \mb in
comparison to multi-atlas segmentation when using the same number of atlases.
The actual benefit in using \mb is consistency of the labelling regardless of
atlas or template choice. This is an important consideration that few have
touched on previously. The 10-fold Monte Carlo leave-one-out cross-validation that we present
in Experiment 1 is amongst one of the most stringent performed in the
multi-atlas/segmentation literature.  To the best of our knowledge, other groups
using ADNI data for validation do at most a single round of leave-one-out-validation
(Table \ref{tab:other-methods}), aside from \citep{Wang2011} who perform a 
similar validation as us. The thoroughness of our validation suggests that
our results are reflective of a true average over the choice of parameter
settings and are independent of atlas or template choice (provided the input
atlases are properly segmented).

% Validation of ADNI in other works
% Chupin 2009: LOOCV, 1-fold validation
% Heckemann 2011: QC of segmentations, comparison across field strengths, 
%                 agreement between two different atlas sets
% Coupe 2011e: LOOCV
% Lotjonen 2010: LOOCV, 1-fold validation
% Leung 2010: 1-fold, QC, classification
% Wolz 2010: 1-fold over different leap settings, classification
% Collins 2010: 1-fold validation, QC
% Morra 2008: LOOCV with 21 atlases, 2 manual raters
% Wang 2011: 10 rounds of MCCV with 20 atlases and 20 subjects, chosen from 
%            a pool of 139 images
